---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# cSEM.DGP: Generate data for structural equation models

[![CRAN status](https://www.r-pkg.org/badges/version/cSEM.DGP)](https://cran.r-project.org/package=cSEM.DGP)
[![Build Status](https://travis-ci.com/M-E-Rademaker/cSEM.DGP.svg?branch=master)](https://travis-ci.com/M-E-Rademaker/cSEM.DGP)

Please report bugs to [the package developer](mailto:manuel.rademaker@uni-wuerzburg.de).

## Installation:
```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("M-E-Rademaker/cSEM.DGP")
```
Note: requires a development version of [cSEM](https://github.com/M-E-Rademaker/cSEM) 
      that includes commits done before 12.09.2019.
      
If you have an older version. Run:
```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("M-E-Rademaker/cSEM")
```

## Purpose
Generate data for structural equation models including up to 8 constructs. Generation
is based on parameter values given in [lavaan model syntax](http://lavaan.ugent.be/tutorial/syntax1.html).

In addition to supplying numeric values, variable values for parameters are allowed.
To achieve this, the package makes use of [lavaan](http://lavaan.ugent.be/)'s 
labeling capabilities. 
Users may replace a given parameter in, i.e. the structural model
by a symbolic name and assign a vector of values to that name. These values will be used
to generate data for all possible combinations of these values with the remaining fixed parameters.

## Examples
### Without variable paramters
Simply write your model as usual in lavaan model syntax. Add a fixed numeric value 
for each parameter. Note, currently you must either set all paramters or none. The type
of output can be choosen. Either a data.frame (`return_type = "data.frame"`, the default),
a numeric matrix (`return.type = "matrix"`), or a correlation matrix (`return.type = "cor"`).
```{r, message=FALSE, warning=FALSE}
require(cSEM.DGP)

model <- "
# Structural model
eta2 ~ 0.6*eta1
eta3 ~ 0.4*eta1 + 0.35*eta2

# Measurement model
eta1 =~ 0.8*y11 + 0.9*y12 + 0.8*y13
eta2 =~ 0.7*y21 + 0.7*y22 + 0.9*y23
eta3 =~ 0.9*y31 + 0.8*y32 + 0.7*y33
"

dat <- generateData(model, .return_type = "cor")
dat
```

### Including variable paramters

To include variable paramters:

1. Replace a numeric value by a character string of your choice.
2. Supply a named vector of values to `generateData()` where the name corresponds to
   the character string of your choice.

Now, data for all possible combinations of these values with the remaining fixed
parameters will be generated

```{r}
# We want to vary the path coefficient between eta2 and eta1 and 
# the first loading of eta1.
model <- "
# Structural model
eta2 ~ gamma1*eta1
eta3 ~ gamma2*eta1 + 0.35*eta2

# Measurement model
eta1 =~ lambda*y11 + 0.9*y12 + 0.8*y13
eta2 =~ 0.7*y21 + 0.7*y22 + 0.9*y23
eta3 =~ 0.9*y31 + 0.8*y32 + 0.7*y33
"

dat <- generateData(model, .return_type = "cor", 
                    lambda = c(0.8, 0.9),
                    gamma1 = c(0.2, 0.3),
                    gamma2 = c(0, 0.2, 0.3),
                    .handle_negative_definite = "drop"
                    )
dat
```
